---
title: 'Case Study #1'
author: "Aniruddha Mukherjee"
date: "3/16/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Setup

#### Libraries Used in this Case Study

Below are the libraries used for this case study.
```{r, echo = TRUE}
library(ggplot2)
```

#### Reading the Data

Below is a sample output of the data that we are given. Note that the column names have been added by us.
```{r, echo = TRUE}
cdi = read.table("CDI.txt", na.strings=".", header=FALSE,
col.names = c("ID", "County", "State", "Land_Area", "Total_Pop", "PercentPop18_24", "PercentPop65_",
              "Num_Active_Phy", "Num_Hosp_Beds", "Total_Serious_crimes", "Percent_HS",
              "Percent_Bachelors", "Percent_Poverty", "Percent_Unemployment", "Per_Capita_Income", 
              "Total_Personal_Income", "Geographic_region"))
head(cdi)
```

To verify that the data that we have read has the correct dimensions, they are displayed below. The data should have $440$ rows and $17$ columns.
```{r, echo = TRUE}
dim(cdi)
```

### Summary of Variables

A display of the summary statistics of some of the quantitative variables are shown below. Along with the summary statistcs, a histogram and an appropriate scatter plot for a few variables are also shown below. This section aims to get the reader a better understanding of the variables that are present in this report. 

The summary statistics of the percent of the population that is aged $18$ to $24$ is displayed below.
```{r, echo = TRUE}
cdi.PercentPop18_24 = summary(cdi$PercentPop18_24)
cdi.PercentPop18_24
```



The summary statistics of the percent of the population that is aged $65$ or older is displayed below.
```{r, echo = TRUE}
cdi.PercentPop65_ = summary(cdi$PercentPop65_)
cdi.PercentPop65_
```



The summary statistic of the number of active nonfederal physicians during $1990$ is shown below is displayed below.
```{r, echo = TRUE}
cdi.Num_Active_Phy = summary(cdi$Num_Active_Phy)
cdi.Num_Active_Phy
```



A histogram of the number of active nonfederal physicians during $1990$ in each county is shown below.
```{r, echo = TRUE}
cdi.id_phy = as.data.frame(cdi[c("ID", "Num_Active_Phy")])
barplot(height = cdi.id_phy$Num_Active_Phy, names = cdi.id_phy$ID, xlab = "ID of Counties",
        ylab = "Number of Physicians", main = "Number of Physicians by County", 
        col = c("red", "green", "blue", "yellow"), border = c("red", "green", "blue", "yellow"))
```

Based on the histogram above, there seems to be a few outliers in the number of Physicians. To explore this idea, a scatter plot is shown below.



A scatter plot of the number of active nonfederal physicians during $1990$ in each county is shown below.
```{r, echo = TRUE}
ggplot(cdi.id_phy, aes(x = ID, y = Num_Active_Phy)) + 
  geom_point( size = 1, shape = 23, color = "red" ) + 
              ylab("Number of Active Physicians") +  xlab("ID of Counties") +
              ggtitle("Scatter Plot of the Number of Physicians in Each County") + 
  theme_update(plot.title = element_text(hjust = 0.5)) + 
  theme_classic()
```



Based on the scatter plot above, two conties seem to have a much higher number of physicians than all other counties.
The two counties with a higher number of physicians are shown below.
```{r, echo = TRUE}
cdi.df <- as.data.frame(cdi)
head(cdi.df[ order(-cdi.df$Num_Active_Phy), ], 2)
```

As seen in the output above, the two counties with the greatest number of physicians in 1990 are **Los Angeles** county and **Cook** county.



The summary statistic of the number of beds, cribs and bassinets during $1990$ is shown below is displayed below.
```{r, echo = TRUE}
cdi.Num_Hosp_Beds = summary(cdi$Num_Hosp_Beds)
cdi.Num_Hosp_Beds
```



A histogram of the number of beds, cribs and bassinets during $1990$ in each county is shown below.
```{r, echo = TRUE}
cdi.id_hos = as.data.frame(cdi[c("ID", "Num_Hosp_Beds")])
barplot(height = cdi.id_hos$Num_Hosp_Beds, names = cdi.id_hos$ID, xlab = "ID of Counties",
        ylab = "Number of Hospital Beds", main = "Number of Hospital Beds by County", 
        col = c("red", "green", "blue", "yellow"), border = c("red", "green", "blue", "yellow"))
```



Based on the histogram above, there seems to be a few outliers in the number of hospital beds. To explore this idea, a scatter plot is shown below.
```{r, echo = TRUE}
ggplot(cdi.id_hos, aes(x = ID, y = Num_Hosp_Beds)) + 
  geom_point( size = 1, shape = 23, color = "red") + 
              ylab("Number of Hospital Beds") +  xlab("ID of Counties") +
              ggtitle("Scatter Plot of the Number of Hospital Beds in Each County") + 
  theme_update(plot.title = element_text(hjust = 0.5)) + 
  theme_classic()
```



Based on the scatter plot above, three conties seem to have a much higher number of hospital beds than all other counties.
The three counties with a higher number of physicians are shown below.
```{r, echo = TRUE}
head(cdi.df[ order(-cdi.df$Num_Hosp_Beds), ], 3)
```

The three counties with the most number of hospital beds are **Los Angeles** county, **Cook** county and **Harris** county. 



The summary statistic of the total number of serious crimes in 1990 as reported by law enforcement agencies is displayed below.
```{r, echo = TRUE}
cdi.Total_Serious_crimes = summary(cdi$Total_Serious_crimes)
cdi.Total_Serious_crimes
```



The summary statistic of the percent of the population that completed $12$ or more years of education (high school education) is displayed below.
```{r, echo = TRUE}
cdi.Percent_HS = summary(cdi$Percent_HS)
cdi.Percent_HS
```



The summary statistic of the percent of the adult population with bachelor's degrees is displayed below.
```{r, echo = TRUE}
cdi.Percent_Bachelors = summary(cdi$Percent_Bachelors)
cdi.Percent_Bachelors
```



The summary statistic of the percent of CDI population with income below poverty level is displayed below.
```{r, echo = TRUE}
cdi.Percent_Poverty = summary(cdi$Percent_Poverty)
cdi.Percent_Poverty
```



The summary statistic of the per capita income of 1990 CDI population (in U.S. dollars) is displayed below.
```{r, echo = TRUE}
cdi.Per_Capita_Income = summary(cdi$Per_Capita_Income)
cdi.Per_Capita_Income
```



The summary statistic of the total personal income of 1990 CDI population (in millions of U.S. dollars) is displayed below.
```{r, echo = TRUE}
cdi.Total_Personal_Income = summary(cdi$Total_Personal_Income)
cdi.Total_Personal_Income
```



The summary statistics of the percent of 1990 CDI labor force that is unemployed is displayed below.
```{r, echo = TRUE}
cdi.Percent_Unemployment = summary(cdi$Percent_Unemployment)
cdi.Percent_Unemployment
```



Histogram of Geographic Region is shown below.
```{r, echo = TRUE}
ggplot(cdi, aes(x = Geographic_region, fill = Geographic_region)) +
  geom_bar(position = "identity", bins = 4, fill = c("red", "green", "blue", "yellow")) +
  labs(title = "Distribution of Geographic Region", x = "Geographic Region") + 
  scale_x_discrete(limits=c("NE", "NC", "S", "W")) + 
  theme_minimal() + ggeasy::easy_center_title()
```


## Statistical Analysis

```{r, echo = TRUE}
cdi = read.table("CDI.txt", na.strings=".", header=FALSE,
col.names = c("ID", "County", "State", "Land_Area", "Total_Pop", "PercentPop18_24", "PercentPop65_",
              "Num_Active_Phy", "Num_Hosp_Beds", "Total_Serious_crimes", "Percent_HS",
              "Percent_Bachelors", "Percent_Poverty", "Percent_Unemployment", "Per_Capita_Income", 
              "Total_Personal_Income", "Geographic_region"))
```

Many factors can have an effect in the number of active physicians in a county. Based on the county Demographic Information (CDI) that we are provided with, we hypothesize that the total estimated population in a county in 1990, the total land area of the county and the total personal income in 1990 of the residents of a county best predict the number of active physicians in a county.The rationale behind these three variables is as follows: Higher populations should require more medical practitioners to account for the increased population, greater land area should increase the number of medical facilities, and wealthier counties seek out medical practitioners more often than less affluent counties. Based on the hypothesized effect that total population, land area and total personal income have, the multiple linear regression model is shown below: 

$$
  Y_i = \beta_{0}+ \beta_1X_{i1}+\beta_2X_{i2}+\beta_3X_{i3}+\varepsilon_i, \quad i = 1, \ldots, n
$$

In the general form multiple linear regression above, $\beta_0$ represents the intercept of our regression model. Similarly, $\beta_1$, $\beta_2$, and $\beta_3$ represent the change in number of physicians that each predictor will have given one unit of increase in total population, land area and total personal income, respectively. 

**Output 1.1**, shown below, displays summary of fitting the desired linear regression model with the three initial hypothesized variables. Among other relevant figures, the values for $\hat\beta_0$, $\hat\beta_1$, $\hat\beta_2$, and $\hat\beta_3$ are shown below.

```{r cdimodel0, echo = FALSE}
cdimodel0 = lm(Num_Active_Phy ~ Total_Pop + Land_Area + Total_Personal_Income, cdi)
cdimodel0.summary = summary(cdimodel0)
print(cdimodel0.summary)
```

Based on **output 1.1**, as shown above we have the following, 
\begin{align}
  \begin{aligned}
    \hat\beta_0 &= -13.32     \\
    \hat\beta_1 &=  0.0008366 \\
    \hat\beta_2 &= -0.06552   \\
    \hat\beta_3 &=  0.09413   \\
  \end{aligned}
\end{align}

Therefore, based on the output above, we have the following estimated multiple regression line:

$$
  \hat{Y_i} = -13.32  + 0.0008366 \cdot X_{i1} - 0.06552 \cdot X_{i2}+0.09413 \cdot X_{i3}
$$

Based on **output 1.1**, the $R^2$ is $0.9026$, this means that $90.26\%$ of the variation in number of physicians can be explained by our three predictors: total population, land area, and total personal income. Similary, based on **output 1.1**, the p-value of each of the estimated coefficients for our regression model is less than a significance level of $0.05$, i.e. $\alpha = 0.05$. Therefore all three predictors are signifincant. 

To increase the accuracy of the model, we hypothesize that number of hospital beds, total serious crimes, and percent of population $65$ or older will significantly improve our ability to predict the number of physicians in a given county. By adding three more explanatory variables, the multiple linear regression model is as follows:

$$
  Y_i = \beta_{0}+ \beta_1X_{i1}+\beta_2X_{i2}+\beta_3X_{i3}+\beta_4X_{i4}+\beta_5X_{i5}+\beta_6X_{i6}+\varepsilon_i, \quad i = 1, \ldots, n
$$

In the general form multiple linear regression above, $\beta_0$ represents the intercept of our regression model. The representation of $\beta_1$, $\beta_2$, and $\beta_3$ are the same as the representation in our previous model.Similarly, $\beta_5$ and $\beta_6$ represent change in the number of physicians for a unit increase in the total number of hospital beds and total serious crimes, respectively. Finally, $\beta_4$ represents a $0.01$ increase in number of physicians given a $1\%$ increase in population aged $65$ or older.


**Output 1.2**, shown below, displays summary of fitting the desired linear regression model with the six hypothesized variables. Among other relevant figures, the values for $\hat\beta_0$, $\hat\beta_1$, $\hat\beta_2$, $\hat\beta_3$, $\hat\beta_4$, $\hat\beta_5$, and $\hat\beta_6$ are shown below.

```{r cdimodel01, echo = FALSE}
cdimodel01 = lm(Num_Active_Phy ~ Total_Pop + Land_Area + Total_Personal_Income + PercentPop65_ + Num_Hosp_Beds + Total_Serious_crimes, cdi)
cdimodel01.summary = summary(cdimodel01)
print(cdimodel01.summary)
```

Based on **output 1.2**, as shown above we have the following, 
\begin{align}
  \begin{aligned}
    \hat\beta_0 &=   79.69     \\
    \hat\beta_1 &=  -0.001916 \\
    \hat\beta_2 &=   0.01964  \\
    \hat\beta_3 &=   0.01403   \\
    \hat\beta_4 &=  -15.77    \\
    \hat\beta_5 &=   0.5161    \\
    \hat\beta_6 &=  -0.000888 \\
  \end{aligned}
\end{align}



Therefore, based on the output above, we have the following estimated multiple regression line:

$$
  \hat{Y_i} = 79.69-0.001916\cdot X_{i1}+0.01964\cdot X_{i2}+0.01403 \cdot X_{i3}-15.77\cdot X_{i4}+0.5161\cdot X_{i5}-0.000888\cdot X_{i6}
$$

Based on **output 1.2**, the $R^2$ is $0.9568$, this means that $95.68\%$ of the variation in number of physicians can be explained by our six predictors: total population, land area, total personal income, percent of the county population aged $65$ or older, number of hospital beds, and total serious crimes. Similary, based on **output 1.2**, not all of the variables have a p-value less than a significance level of $0.05$, i.e. $\alpha = 0.05$. Therefore not all of the variables are significant. The two variables that have a p-value greater than a significance level of $5\%$ are **Land_Area** and **Total_Serious_Crimes**. 

By dropping the two variables with a significance level below $0.05$ the following updated multiple linear regression model is obtained:

$$
  Y_i = \beta_{0}+ \beta_1X_{i1}+\beta_2X_{i2}+\beta_3X_{i3}+\beta_4X_{i4}+\varepsilon_i, \quad i = 1, \ldots, n
$$

In the general form multiple linear regression above, $\beta_0$ represents the intercept of our regression model. The representation of $\beta_1$ is the change in the number of physicians for a unit increase in total county population. The representation of $\beta_2$  is the change in the number of physicians for a unit increase in total personal income. The representation of $\beta_3$ is  a $0.01$ increase in number of physicians given a $1\%$ increase in population aged $65$ or older. Finally, $\beta_4$ is the change in the number of physicians in a unit change in the number of hospital beds.

**Output 1.3**, shown below, displays summary of fitting the desired linear regression model with the four hypothesized variables. Among other relevant figures, the values for $\hat\beta_0$, $\hat\beta_1$, $\hat\beta_2$, $\hat\beta_3$, and $\hat\beta_4$ are shown below.

```{r, echo = TRUE}
cdimodel02 = lm(Num_Active_Phy ~ Total_Pop + Total_Personal_Income + PercentPop65_ + Num_Hosp_Beds, cdi)
cdimodel02.summary = summary(cdimodel02)
print(cdimodel02.summary)
```


Based on **output 1.2**, as shown above we have the following, 
\begin{align}
  \begin{aligned}
    \hat\beta_0 &=   92.87    \\
    \hat\beta_1 &=  -0.001930 \\
    \hat\beta_2 &=   0.1402   \\
    \hat\beta_3 &=  -14.89    \\
    \hat\beta_4 &=   0.5015   \\
  \end{aligned}
\end{align}

Therefore, based on the output above, we have the following estimated multiple regression line:

$$
  \hat{Y_i} = 92.87-0.001930\cdot X_{i1}+0.1402\cdot X_{i2}-14.89 \cdot X_{i3}+0.5015\cdot X_{i4}
$$

Based on **output 1.3**, the $R^2$ is $0.9563$, this means that $95.63\%$ of the variation in number of physicians can be explained by four predictors: total population, total personal income, percent of the county population aged $65$ or older, and number of hospital beds. Similary, based on **output 1.3**, the p-value of each of the estimated coefficients for our regression model is less than a significance level of $0.05$, i.e. $\alpha = 0.05$. Therefore all four predictors are signifincant.

The model assumptions for a multiple linear regression model are the following, 

+ **Constant Variance**

+ **Normality**

+ **Uncorrelated Errors** 


To check for **constant variance**, the residuals against the fitted values are shown below in **plot 1.1**.
```{r, echo = TRUE}
plot(cdimodel02, which = 1)
```

The scatter plot, **plot 1.1**, shows that the model suffers from heteroscedacity. 

```{r, echo = FALSE}
cdi.log <- log(cdi[,-c(1, 2, 3)])
cdimodel02_log = lm(Num_Active_Phy ~ Total_Pop + Total_Personal_Income + PercentPop65_ + Num_Hosp_Beds, cdi.log)
cdimodel02_log.summary = summary(cdimodel02_log)
print(cdimodel02_log.summary)
```


```{r, echo = FALSE}
plot(cdimodel02_log, which = 1)
```

The plot above has homoscedacity.


```{r qq, echo = FALSE}
plot(cdimodel02_log, which = 2)
```

QQ plot looks about normal.



```{r, echp = FALSE}
library(lmtest)
```







<!-- Notice that the model shown in **output 1.1** and in the **output 1.3** both significant coefficients and their corresponding $R^2$ values are similar. An ANOVA test is performed to decide which model should be used in predicting the number of physicians in a given county with the following hypothesis tests. -->

<!-- \begin{align} -->
<!--   \begin{aligned} -->
<!--     H_0 &= \text{ The reduced model is } \\ -->
<!--     H_{\alpha} &= \text{ The full model is required } \\ -->
<!--   \end{aligned} -->
<!-- \end{align} -->

<!-- Please note that reduced model refers to the  -->



















