---
title: 'Case Study #1'
author: "Aniruddha Mukherjee"
date: "3/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Setup

#### Libraries Used in this Case Study

Below are the libraries used for this case study.
```{r, echo = TRUE}
library(ggplot2)
```

#### Reading the Data

Below is a sample output of the data that we are given. Note that the column names have been added by us.
```{r, echo = TRUE}
cdi = read.table("CDI.txt", na.strings=".", header=FALSE,
col.names = c("ID", "County", "State", "Land_Area", "Total_Pop", "PercentPop18_24", "PercentPop65_",
              "Num_Active_Phy", "Num_Hosp_Beds", "Total_Serious_crimes", "Percent_HS",
              "Percent_Bachelors", "Percent_Poverty", "Percent_Unemployment", "Per_Capita_Income", 
              "Total_Personal_Income", "Geographic_region"))
head(cdi)
```

To verify that the data that we have read has the correct dimensions, they are displayed below. The data should have $440$ rows and $17$ columns.
```{r, echo = TRUE}
dim(cdi)
```

### Summary of Variables

A display of the summary statistics of the quantitative variables are shown below

```{r, echo = TRUE}

# Summary Statistics
# cdi.PercentPop18_24 = summary(cdi$PercentPop18_24)
# cdi.PercentPop65_ = summary(cdi$PercentPop65_)
# cdi.Num_Active_Phy = summary(cdi$Num_Active_Phy)
# cdi.Num_Hosp_Beds = summary(cdi$Num_Hosp_Beds)
# cdi.Total_Serious_crimes = summary(cdi$Total_Serious_crimes)
# cdi.Percent_HS = summary(cdi$Percent_HS)
# cdi.Percent_Bachelors = summary(cdi$Percent_Bachelors)
# cdi.Percent_Poverty = summary(cdi$Percent_Poverty)
# cdi.Percent_Unemployment = summary(cdi$Percent_Unemployment)
# cdi.Per_Capita_Income = summary(cdi$Per_Capita_Income)
# cdi.Total_Personal_Income = summary(cdi$Total_Personal_Income)
# 
# # Histogram for Geographic Region
# hist(cdi$Geographic_region)
# 
# # Linear Regression with Suggested Variables ?
# cdimodel0 = lm(Num_Active_Phy ~ Total_Pop + Land_Area + Total_Personal_Income, cdi)
# 
# cdimodel0.summary = summary(cdimodel0)
# 
# print(cdimodel0.summary) # Al variables seem significant especially Total_Personal_Income

# Checking for model assumptions

  # Check for Linearity
  #

  # Check for Homoscdedacity
  # residual plot

  # Check for Independence

  # Check for Normality


# Linear Regression
# cdimodel1 = lm(Num_Active_Phy ~ Total_Serious_crimes + Per_Capita_Income + Num_Hosp_Beds, cdi)

# cdimodel1.summary = summary(cdimodel1)

# print(cdimodel1.summary) # Total_Serious_crimes p-val (0.0636) > 0.05

# cdimodel2 = lm(Num_Active_Phy ~ Per_Capita_Income + Num_Hosp_Beds, cdi)

# cdimodel2.summary = summary(cdimodel2)

# print(cdimodel2.summary)

# anova_mod1_mod2 = anova(cdimodel1, cdimodel2)

# print(anova_mod1_mod2) # Fail to reject null, proceed with cdimodel2

# Analysis with
```




































